A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-05 paper.tex
edits in sect. 4

% functions
\newcommand{\meet}[2]{\ensuremath{#1 \sqcap #2}}
\newcommand{\denote}[1]{\ensuremath{\llbracket #1 \rrbracket}}
\newcommand{\proxy}[2]{\ensuremath{\mathrm{proxy}(#1,#2)}}
\newcommand{\applyCast}[2]{\ensuremath{\mathrm{applyCast}(#1,#2)}}
\newcommand{\project}[4]{\ensuremath{\mathrm{project}(#1,#2,#3,#4)}}
\newcommand{\composegap}[3]{\ensuremath{#1 \fatsemi #2 \fatsemi #3}}
\newcommand{\mkhead}[2]{\ensuremath{\mathrm{mkProj}(#1,#2)}}
\newcommand{\mktail}[2]{\ensuremath{\mathrm{mkInj}(#1,#2)}}
\newcommand{\mkgap}[3]{\ensuremath{\mathrm{mkGap}(#1,#2,#3)}}
\newcommand{\compose}[2]{\ensuremath{#1 \fatsemi #2}}
\newcommand{\translate}[1]{\ensuremath{\lceil#1\rceil}}
\newcommand{\castToDyn}[2]{\ensuremath{\mathrm{castToDyn}(#1,#2)}}
\newcommand{\castFromDyn}[2]{\ensuremath{\mathrm{castFromDyn}(#1,#2)}}
\newcommand{\dom}[1]{\ensuremath{\mathrm{dom}(#1)}}
\newcommand{\cod}[1]{\ensuremath{\mathrm{cod}(#1)}}
\newcommand{\fst}[1]{\ensuremath{\mathrm{fst}(#1)}}
\newcommand{\snd}[1]{\ensuremath{\mathrm{snd}(#1)}}
\newcommand{\id}[1]{\ensuremath{\mathrm{id}(#1)}}
\newcommand{\idm}[1]{\ensuremath{\mathrm{id}_m(#1)}}
\newcommand{\idc}[1]{\ensuremath{\mathrm{id}_c(#1)}}
\newcommand{\mkCont}[1]{\ensuremath{\mathrm{mkK}(#1)}}
\newcommand{\extCont}[2]{\ensuremath{\mathrm{extK}(#1,#2)}}
% \newcommand{\applyCont}[2]{\ensuremath{\langle\langle#1,#2\rangle\rangle}}
\newcommand{\applyCont}[2]{\ensuremath{\mathrm{applyK}(#1,#2)}}

@@

\begin{figure}
	\fbox{\reducetoR{B}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreduceto{\sexpr{\elam{x}{S}{e}}{E}{K}}{\scont{\vfunc{x}{S}{e}{E}}{K}}{}
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FappXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\scont{e_1}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FconsXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\efst{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Ffst}{K}}}{}
	\figreduceto{\sexpr{\esnd{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fsnd}{K}}}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}}{
		\applyCast{v}{\ccast{S}{l}{T}} \; \mbind \; \lambda u. \scont{u}{K}
	}{}
	\figreduceto{\scont{v_1}{\Kcons{\FappXO{e_2}{E}}{K}}}{\sexpr{e_2}{E}{
			\Kcons{\FappOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v}{\Kcons{\FappOX{\vfunc{x}{S}{e}{E}}}{K}}}{
		\sexpr{e}{E[x:=v]}{K}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\FappOX{(\vfuncc{u}{S_1}{T_1}{l}{S_2}{T_2})}}{K}}}{
		\scont{v}{
			\Kcons{\Fcast{\ccast{S_2}{\bar{l}}{S_1}}}{
				\Kcons{\FappOX{u}}{
					\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}}
	}{}
	\figreduceto{\scont{\vtrue}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_2}{E}{K}}{}
	\figreduceto{\scont{\vfalse}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_3}{E}{K}}{}
	\figreduceto{\scont{v_1}{\Kcons{\FconsXO{e_2}{E}}{K}}}{
		\sexpr{e_2}{E}{\Kcons{\FconsOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v_2}{\Kcons{\FconsOX{v_1}}{K}}}{
		\scont{\vcons{v_1}{v_2}}{K}
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Ffst}{K}}}{\scont{v_1}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Ffst}{K}}}{
		\scont{u}{
			\Kcons{\Ffst}{
				\Kcons{\Fcast{\ccast{S_1}{l}{S_2}}}{K}}}	
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fsnd}{K}}}{\scont{v_2}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fsnd}{K}}}{
		\scont{u}{
			\Kcons{\Fsnd}{
				\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}	
	}{}
	\figreduceto{\scont{v}{\Knil}}{\shalt{v}}{}
	\end{array}
	\]

@@

\begin{figure}
\[
\begin{array}{lrcl}
\stxrule{I,J}{Injectable types}{P}
\end{array}
\]

@@

\begin{figure}
	\[
	\begin{array}{lrcl}
	\stxrule{c,d}{Runtime casts}{\text{a data type defined by the 
	instance of Cast ADT}}
	\stxrule{v,u}{Values}{
		\vdyn{I}{v}  \mid
		\vtrue       \mid
		\vfalse                     
	}
	\stxrulecont{
		\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c}{c}} \mid
		\ecast{\vcons{v}{v}}{\Pprod{c}{c}}
	}
	\stxrule{I,J}{Injectable types (\LD)}{P}
	\stxrule{I,J}{Injectable types (\LUD)}{G}
	\stxrule{E}{Environments}{
		\text{a partial function}\;\{\langle x , v \rangle , \dots \}
	}
	\stxrule{F}{Frames}{
		\FappXO{e}{E}  \mid
		\FappOX{v}     \mid
		\Fif{e}{e}{E}
	}
	\stxrulecont{
		\Ffst \mid
		\Fsnd \mid
		\FconsXO{e}{E}  \mid
		\FconsOX{v}
	}
	\stxrule{K}{Continuations}{
		\Kcons{\Fcast{c}}{k}
	}
	\stxrule{k}{Pre-continuations}{
		\Knil         \mid
		\Kcons{F}{K}
	}
	\stxrule{s}{States}{
		\sexpr{e}{E}{K}  \mid
		\scont{v}{K}     \mid
		\shalt{v}        \mid
		\error{l}
	}
	\stxrule{r}{Cast results}{
		v \mid
		\error{l}
	}
	\end{array}
	\]

@@

\figref{fig:SEMachine-state} defines the syntax of the space-efficient
CEK machine \SMachine{C}.  The key differences between \SMachine{C}
and \CMachine{B} are the places where casts can accumulate, that is,
in the definitions of values and continuations.

@@

Before moving to the transition relation of \SMachine{C}, let's define two 
helper functions that manipulate continuations.
%
The function $\mathit{mkCont}$ construct a continuation from a pre-continuation 
by topping the pre-continuation with an identity cast. 
%
The function $\mathit{extCont}$ composes $c$ with the cast at the top of $K$ by 
using the composition operator 
$\fatsemi$.
\[
\begin{array}{rclr}
\figequalto{\mkCont{k}}{\Kcons{\Fcast{\id{S}}}{k}}{\sidecond{\casttyping{k}{S}{T}}}
\figequalto{\extCont{c}{\Kcons{\Fcast{d}}{k}}}{
	\Kcons{\Fcast{(\compose{c}{d})}}{k}
}{}
\end{array}
\]

\begin{figure}
	\fbox{\reducetoS{C}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreducetoH{
		\sexpr{\elam{x}{S}{e}}{E}{K}
	}{
		\scont{\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{id(S)}{id(T)}}}{K}
	}{}
	& &
	\highlight{
	\sidecond{
		\expressiontyping{\Gamma,S}{e}{T}
		\text{ and }
		\Gamma \vdash E}
	}
	\\
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\FappXO{e_2}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\FconsXO{e_2}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\efst{e}}{E}{K}}{
		\sexpr{e}{E}{\mkCont{\Kcons{\Ffst}{K}}}
	}{}
	\figreduceto{\sexpr{\esnd{e}}{E}{K}}{
		\sexpr{e}{E}{\mkCont{\Kcons{\Fsnd}{K}}}
	}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\extCont{\translate{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{c}}{k}}}{
		\denote{c}(v) \; \mbind \; \lambda u. \applyCont{u}{K}
	}{}
	\end{array}
	\]

@@

\figref{fig:semachine-dynamics} defines the transition relation 
$\reducetoS{C}{s}{s}$. They rely on functions provided 
by $C$ to work with casts.
%
To construct a function value, the machine fills the casts with identities.
%
When evaluating a cast expression, the machine extends the continuation with 
\translate{\ccast{S}{l}{T}}.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the cast application fails, the machine halts with the blame 
label from the failure. Otherwise, the machine handles the pre-continuation 
with the function $\applyCont{v}{k}$.
%
To apply a function, the machine applies the domain cast $c$ to the 
argument $v$ and evaluates the function body. The codomain cast $d$ is merged 
into the continuation with $\mathrm{extK}$.
%
When the machine constructs a pair, it fills the casts with identities, just 
like how it does for function values. 
%
To take out the first part of a pair, the machine applies the first pending 
cast on the first value and returns the result to the continuation.
%
Taking out the second part of a pair is similar.

@@

\begin{figure}

@@

\begin{figure}
	Syntax of Hypercoercions
	\[
	\begin{array}{lrcl}
	\stxrule{c}{Hypercoercions}{
		\hcci           \mid
		\hccc{h}{m}{t}
	}
	\stxrule{h}{Heads}{
		\hche     \mid
		\hchp{I}{l}
	}
	\stxrule{m}{Middles}{
		\hcmbool       \mid
		\hcmfunc{c}{c} \mid
		\hcmprod{c}{c} \mid
		\hcmfail{l}    
	}
	\stxrule{t}{Tails}{
		\hcte     \mid
		\hcti{I}
	}
	\end{array}
	\]

@@

\begin{figure} %% composition 1
\[
\begin{array}{lrcl}
\stxrule{g}{Gaps}{
	\gnone           \mid
	\gsome{S}{l}{T}
}
\end{array}
\]

@@

\begin{figure}
	\fbox{$\translate{\ccast{S}{l}{T}} = c$} Translation of \LD\ hypercoercions
	\[
	\begin{array}{rclr}
	\figequalto{\translate{\ccast{S}{l}{T}}}{\composegap{id(S)}{(\gsome{S}{l}{T})}{id(T)}}{}
	\end{array}
	\]
	\caption{Translatin of \LD\ Hypercoercions}
	\label{fig:LazyD-Hypercoercions-translation}
\end{figure}

@@

\begin{figure}
\fbox{$\denote{c}(v)=r$}
\[
\begin{array}{rclr}
\figequalto{\denote{\hcci}(v)}{v}{}
\figequalto{\denote{\hccc{h}{m}{t}}(v)}{
	(\denote{h}_h(v) \mbind \denote{m}_m)
	\mbind \denote{t}_t
}{}
\end{array}
\]

@@

\begin{figure}
	\fbox{$c \fatsemi c = c$}
	\[
	\begin{array}{rclcl}
		\hcci          
		& \fatsemi & d
		& = & d \\
		\hccc{h}{m}{t}
		& \fatsemi & \hcci
		& = & \hccc{h}{m}{t} \\
		\hccc{h_1}{m_1}{t_1}
		& \fatsemi & \hccc{h_2}{m_2}{t_2}
		& = & \hccc{h_1}{(m_1 \fatsemi t_1 \fatsemi h_2 \fatsemi m_2)}{t_2}
	\end{array}
	\]

@@

\begin{figure}
	\fbox{$\translate{\ccast{S}{l}{T}} = c$}
	\[
	\begin{array}{rcll}
	\figequalto{\translate{\ccast{\Tdyn}{l}{\Tdyn}}}{\hcci}{}
	\figequalto{\translate{\ccast{\Tdyn}{l}{Q}}}{\castFromDyn{l}{Q}}{}
	\figequalto{\translate{\ccast{P}{l}{\Tdyn}}}{\castToDyn{P}{l}}{}
	\figequalto{\translate{\ccast{P}{l}{Q}}}{
		\hccc{\hche}{\hcmfail{l}}{\hcte}
	}{\sidecond{\notshallowlyconsistent{P}{Q}}}
	\figequalto{\translate{\ccast{\Pbool}{l}{\Pbool}}}{
		\hccc{\hche}{\hcmbool}{\hcte}
	}{}
	\figequalto{\translate{\ccast{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}{
		\hccc{\hche}{
			\hcmfunc{
				\translate{\ccast{S_2}{l}{S_1}}}{
				\translate{\ccast{T_1}{l}{T_2}}}
		}{\hcte}
	}{}
	\figequalto{\translate{\ccast{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
		\hccc{\hche}{
			\hcmprod{
				\translate{\ccast{S_1}{l}{S_2}}}{
				\translate{\ccast{T_1}{l}{T_2}}}
		}{\hcte}
	}{}
	\end{array}
	\]

@@

\begin{figure}
	\fbox{$\denote{c}_c(v)=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\denote{\hcci}(v)}{v}{}
	\figequalto{\denote{\hccc{h}{m}{t}}(v)}{
		(\denote{h}(v) \mbind \denote{m})
		\mbind \denote{t}
	}{}
	\end{array}
	\]

>>>>

% functions
\newcommand{\meet}[2]{\ensuremath{#1 \sqcap #2}}
\newcommand{\denote}[1]{\ensuremath{\llbracket #1 \rrbracket}}
\newcommand{\proxy}[2]{\ensuremath{\mathrm{proxy}(#1,#2)}}
\newcommand{\applyCast}[2]{\ensuremath{\mathrm{applyCast}(#1,#2)}}
\newcommand{\project}[4]{\ensuremath{\mathrm{project}(#1,#2,#3,#4)}}
\newcommand{\composegap}[3]{\ensuremath{#1 \fatsemi #2 \fatsemi #3}}
\newcommand{\mkhead}[2]{\ensuremath{\mathrm{mkProj}(#1,#2)}}
\newcommand{\mktail}[2]{\ensuremath{\mathrm{mkInj}(#1,#2)}}
\newcommand{\mkgap}[3]{\ensuremath{\mathrm{mkGap}(#1,#2,#3)}}
\newcommand{\compose}[2]{\ensuremath{#1 \fatsemi #2}}
\newcommand{\translate}[1]{\ensuremath{\lceil#1\rceil}}
\newcommand{\castToDyn}[2]{\ensuremath{\mathrm{castToDyn}(#1,#2)}}
\newcommand{\castFromDyn}[2]{\ensuremath{\mathrm{castFromDyn}(#1,#2)}}
\newcommand{\dom}[1]{\ensuremath{\mathrm{dom}(#1)}}
\newcommand{\cod}[1]{\ensuremath{\mathrm{cod}(#1)}}
\newcommand{\fst}[1]{\ensuremath{\mathrm{fst}(#1)}}
\newcommand{\snd}[1]{\ensuremath{\mathrm{snd}(#1)}}
\newcommand{\id}[1]{\ensuremath{\mathrm{id}(#1)}}
\newcommand{\idm}[1]{\ensuremath{\mathrm{id}_m(#1)}}
\newcommand{\idc}[1]{\ensuremath{\mathrm{id}_c(#1)}}
\newcommand{\mkContName}[0]{\ensuremath{\mathrm{mkK}}}
\newcommand{\mkCont}[1]{\ensuremath{\mkContName(#1)}}
\newcommand{\extContName}[0]{\ensuremath{\mathrm{extK}}}
\newcommand{\extCont}[2]{\ensuremath{\extContName(#1,#2)}}
% \newcommand{\applyCont}[2]{\ensuremath{\langle\langle#1,#2\rangle\rangle}}
\newcommand{\applyCont}[2]{\ensuremath{\mathrm{applyK}(#1,#2)}}

@@

\begin{figure}[tp]
	\fbox{\reducetoR{B}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreduceto{\sexpr{\elam{x}{S}{e}}{E}{K}}{\scont{\vfunc{x}{S}{e}{E}}{K}}{}
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FappXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\scont{e_1}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FconsXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\efst{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Ffst}{K}}}{}
	\figreduceto{\sexpr{\esnd{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fsnd}{K}}}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}}{
		\applyCast{v}{\ccast{S}{l}{T}} \; \mbind \; \lambda u. \scont{u}{K}
	}{}
	\figreduceto{\scont{v_1}{\Kcons{\FappXO{e_2}{E}}{K}}}{\sexpr{e_2}{E}{
			\Kcons{\FappOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v}{\Kcons{\FappOX{\vfunc{x}{S}{e}{E}}}{K}}}{
		\sexpr{e}{E[x:=v]}{K}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\FappOX{(\vfuncc{u}{S_1}{T_1}{l}{S_2}{T_2})}}{K}}}{
		\scont{v}{
			\Kcons{\Fcast{\ccast{S_2}{\bar{l}}{S_1}}}{
				\Kcons{\FappOX{u}}{
					\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}}
	}{}
	\figreduceto{\scont{\vtrue}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_2}{E}{K}}{}
	\figreduceto{\scont{\vfalse}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_3}{E}{K}}{}
	\figreduceto{\scont{v_1}{\Kcons{\FconsXO{e_2}{E}}{K}}}{
		\sexpr{e_2}{E}{\Kcons{\FconsOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v_2}{\Kcons{\FconsOX{v_1}}{K}}}{
		\scont{\vcons{v_1}{v_2}}{K}
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Ffst}{K}}}{\scont{v_1}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Ffst}{K}}}{
		\scont{u}{
			\Kcons{\Ffst}{
				\Kcons{\Fcast{\ccast{S_1}{l}{S_2}}}{K}}}	
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fsnd}{K}}}{\scont{v_2}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fsnd}{K}}}{
		\scont{u}{
			\Kcons{\Fsnd}{
				\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}	
	}{}
	\figreduceto{\scont{v}{\Knil}}{\shalt{v}}{}
	\end{array}
	\]

@@

\begin{figure}[tp]
\[
\begin{array}{lrcl}
\stxrule{I,J}{Injectable types}{P}
\end{array}
\]

@@

\begin{figure}[tp]
	\[
	\begin{array}{lrcl}
	\stxrule{c,d}{Runtime casts}{\text{a data type defined by the 
	instance of Cast ADT}}
	\stxrule{v,u}{Values}{
		\vdyn{I}{v}  \mid
		\vtrue       \mid
		\vfalse                     
	}
	\stxrulecont{
		\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{c}{c}} \mid
		\ecast{\vcons{v}{v}}{\Pprod{c}{c}}
	}
	\stxrule{I,J}{Injectable types (\LD)}{P}
	\stxrule{I,J}{Injectable types (\LUD)}{G}
	\stxrule{E}{Environments}{
		\text{a partial function}\;\{\langle x , v \rangle , \dots \}
	}
	\stxrule{F}{Frames}{
		\FappXO{e}{E}  \mid
		\FappOX{v}     \mid
		\Fif{e}{e}{E}
	}
	\stxrulecont{
		\Ffst \mid
		\Fsnd \mid
		\FconsXO{e}{E}  \mid
		\FconsOX{v}
	}
	\stxrule{K}{Continuations}{
		\Kcons{\Fcast{c}}{k}
	}
	\stxrule{k}{Pre-continuations}{
		\Knil         \mid
		\Kcons{F}{K}
	}
	\stxrule{s}{States}{
		\sexpr{e}{E}{K}  \mid
		\scont{v}{K}     \mid
		\shalt{v}        \mid
		\error{l}
	}
	\stxrule{r}{Cast results}{
		v \mid
		\error{l}
	}
	\end{array}
	\]

@@

\figref{fig:SEMachine-state} defines the syntax of the space-efficient
CEK machine \SMachine{C}.  The key differences between \SMachine{C}
and \CMachine{B} are in the places where casts can accumulate, that
is, in the definitions of values and continuations.

@@

Before discussing the transition relation of \SMachine{C}, we define
two helper functions for manipulating continuations.
%
The function $\mkContName$ constructs a continuation from a
pre-continuation by topping it with an identity cast.
\begin{align*}
  \mkCont{k} &= \Kcons{\Fcast{\id{S}}}{k} & \sidecond{\casttyping{k}{S}{T}}
\end{align*}
The function $\extContName$ composes $c$ with the cast at the top of
$K$ by using the composition operator $\fatsemi$.
\begin{align*}
  \extCont{c}{\Kcons{\Fcast{d}}{k}} &= \Kcons{\Fcast{(\compose{c}{d})}}{k}
\end{align*}

\begin{figure}[tp]
	\fbox{\reducetoS{C}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreducetoH{
		\sexpr{\elam{x}{S}{e}}{E}{K}
	}{
		\scont{\ecast{\vfunc{x}{S}{e}{E}}{\Pfunc{id(S)}{id(T)}}}{K}
	}{}
	& &
	\highlight{
	\sidecond{
		\expressiontyping{\Gamma,S}{e}{T}
		\text{ and }
		\Gamma \vdash E}
	}
	\\
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\FappXO{e_2}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\mkCont{\Kcons{\FconsXO{e_2}{E}}{K}}}
	}{}
	\figreduceto{\sexpr{\efst{e}}{E}{K}}{
		\sexpr{e}{E}{\mkCont{\Kcons{\Ffst}{K}}}
	}{}
	\figreduceto{\sexpr{\esnd{e}}{E}{K}}{
		\sexpr{e}{E}{\mkCont{\Kcons{\Fsnd}{K}}}
	}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\extCont{\translate{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{c}}{k}}}{
		\denote{c}(v) \; \mbind \; \lambda u. \applyCont{u}{K}
	}{}
	\end{array}
	\]

@@

\figref{fig:semachine-dynamics} defines the transition relation
$\reducetoS{C}{s}{s}$. It relies on the functions provided by the Cast
ADT $C$.
%
We focus on the transition rules highlighted in red, which involve
casts and the changes needed for space efficiency.
%
The rule for evaluating a function produces a function value, using
identities for the casts on the function value's domain and range.
%
The rule for evaluating a cast expression extends the continuation
with the cast \translate{\ccast{S}{l}{T}}.
%
The rule for returning a value to a continuation first applies the top
cast of the continutation to the value. If the cast application fails,
the machine halts with the blame label from the failure. Otherwise,
the machine handles the pre-continuation with the function
$\applyCont{v}{k}$.
%
To apply a function, the machine applies the domain cast $c$ to the 
argument $v$ and evaluates the function body. The codomain cast $d$ is merged 
into the continuation with $\mathrm{extK}$.
%
When the machine constructs a pair, it fills the casts with identities, just 
like how it does for function values. 
%
To take out the first part of a pair, the machine applies the first pending 
cast on the first value and returns the result to the continuation.
%
Taking out the second part of a pair is similar.

@@

\begin{figure}[tp]

@@

\begin{figure}[tp]
	Syntax of Hypercoercions
	\[
	\begin{array}{lrcl}
	\stxrule{c}{Hypercoercions}{
		\hcci           \mid
		\hccc{h}{m}{t}
	}
	\stxrule{h}{Heads}{
		\hche     \mid
		\hchp{I}{l}
	}
	\stxrule{m}{Middles}{
		\hcmbool       \mid
		\hcmfunc{c}{c} \mid
		\hcmprod{c}{c} \mid
		\hcmfail{l}    
	}
	\stxrule{t}{Tails}{
		\hcte     \mid
		\hcti{I}
	}
	\end{array}
	\]

@@

\begin{figure}[tp] %% composition 1
\[
\begin{array}{lrcl}
\stxrule{g}{Gaps}{
	\gnone           \mid
	\gsome{S}{l}{T}
}
\end{array}
\]

@@

\begin{figure}[tp]
	\fbox{$\translate{\ccast{S}{l}{T}} = c$} Translation of \LD\ hypercoercions
	\[
	\begin{array}{rclr}
	\figequalto{\translate{\ccast{S}{l}{T}}}{\composegap{id(S)}{(\gsome{S}{l}{T})}{id(T)}}{}
	\end{array}
	\]
	\caption{Translatin of \LD\ Hypercoercions}
	\label{fig:LazyD-Hypercoercions-translation}
\end{figure}

@@

\begin{figure}[tp]
\fbox{$\denote{c}(v)=r$}
\[
\begin{array}{rclr}
\figequalto{\denote{\hcci}(v)}{v}{}
\figequalto{\denote{\hccc{h}{m}{t}}(v)}{
	(\denote{h}_h(v) \mbind \denote{m}_m)
	\mbind \denote{t}_t
}{}
\end{array}
\]

@@

\begin{figure}[tp]
	\fbox{$c \fatsemi c = c$}
	\[
	\begin{array}{rclcl}
		\hcci          
		& \fatsemi & d
		& = & d \\
		\hccc{h}{m}{t}
		& \fatsemi & \hcci
		& = & \hccc{h}{m}{t} \\
		\hccc{h_1}{m_1}{t_1}
		& \fatsemi & \hccc{h_2}{m_2}{t_2}
		& = & \hccc{h_1}{(m_1 \fatsemi t_1 \fatsemi h_2 \fatsemi m_2)}{t_2}
	\end{array}
	\]

@@

\begin{figure}[tp]
	\fbox{$\translate{\ccast{S}{l}{T}} = c$}
	\[
	\begin{array}{rcll}
	\figequalto{\translate{\ccast{\Tdyn}{l}{\Tdyn}}}{\hcci}{}
	\figequalto{\translate{\ccast{\Tdyn}{l}{Q}}}{\castFromDyn{l}{Q}}{}
	\figequalto{\translate{\ccast{P}{l}{\Tdyn}}}{\castToDyn{P}{l}}{}
	\figequalto{\translate{\ccast{P}{l}{Q}}}{
		\hccc{\hche}{\hcmfail{l}}{\hcte}
	}{\sidecond{\notshallowlyconsistent{P}{Q}}}
	\figequalto{\translate{\ccast{\Pbool}{l}{\Pbool}}}{
		\hccc{\hche}{\hcmbool}{\hcte}
	}{}
	\figequalto{\translate{\ccast{\Pfunc{S_1}{T_1}}{l}{\Pfunc{S_2}{T_2}}}}{
		\hccc{\hche}{
			\hcmfunc{
				\translate{\ccast{S_2}{l}{S_1}}}{
				\translate{\ccast{T_1}{l}{T_2}}}
		}{\hcte}
	}{}
	\figequalto{\translate{\ccast{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
		\hccc{\hche}{
			\hcmprod{
				\translate{\ccast{S_1}{l}{S_2}}}{
				\translate{\ccast{T_1}{l}{T_2}}}
		}{\hcte}
	}{}
	\end{array}
	\]

@@

\begin{figure}[tp]
	\fbox{$\denote{c}_c(v)=r$}
	\[
	\begin{array}{rclr}
	\figequalto{\denote{\hcci}(v)}{v}{}
	\figequalto{\denote{\hccc{h}{m}{t}}(v)}{
		(\denote{h}(v) \mbind \denote{m})
		\mbind \denote{t}
	}{}
	\end{array}
	\]
